---
title: "Segmentation"
output: html_notebook
---


Algorithms to segment the covid19 cases in India. Just guinea pigging here

```{r}
dailyCases %>%
  mutate(State=str_to_title(State), District=str_to_title(District), City=str_to_title(City)) %>%
  group_by(State, District, City) %>%
  summarise(Cases=sum(Cases)) %>%
  ungroup() ->
  totalCases
```

```{r}
totalCases
```

```{r}
bind_rows(
  totalCases %>%
    group_by(State) %>%
    summarise(Cases=sum(Cases)) %>%
    ungroup() %>%
    mutate(Adjusted=Cases*n(), Region=State) %>%
    arrange(-Adjusted) %>%
    select(Region, Cases, Adjusted),
  
  totalCases %>%
    filter(!is.na(District)) %>%
    group_by(State, District) %>%
    summarise(Cases=sum(Cases)) %>%
    ungroup() %>%
    mutate(Adjusted=Cases*n(), Region=paste(District, State, sep=', ')) %>%
    arrange(-Adjusted) %>%
    select(Region, Cases, Adjusted),
  
  
  totalCases %>%
    filter(!is.na(City)) %>%
    group_by(State, District, City) %>%
    summarise(Cases=sum(Cases)) %>%
    ungroup() %>%
    mutate(Adjusted=Cases*n(), Region=paste(City, District, State, sep=', ')) %>%
    arrange(-Adjusted) %>%
    select(Region, Cases, Adjusted)
) %>%
  arrange(-Adjusted)


```
How do we account for Delhi? It is 1 level deep but it's also a "leaf node".

Leaf or not shouldn't matter for clustering. And in any case, segments will be shown in descending order of size. 

I should forget for this purpose that Maharashtra is much bigger than Delhi - my algo doesn't know this 
This depth comparison is only to compare parents to children. Else doesn't matter. Like should we split Maharashtra into Bombay, Thane, Pune and the rest? 


Let's say we want top 10. Take top 9 clusters (at whatever level). 10th is "rest of India". Now, if there is one that is the subset of another, then the "superset" needs to be changed to "rest of". Now it might slip out of top 9. That's fully fine. 

Only thing is we don't watn a racing condition. Only other input is number of segments


This function is an old and complicated one. Replaced by a simpler one later on 
```{r}

segmentationOld <- function(totalCases, numSegment) {
  
  totalCases %>%
    group_by(Region=State) %>%
    summarise(Cases=sum(Cases)) %>%
    mutate(Level=1) %>%
    bind_rows(
      totalCases %>%
        group_by(Region=District, Parent=State) %>%
        summarise(Cases=sum(Cases)) %>%
        mutate(Level=2)
    ) %>%
    bind_rows(
      totalCases %>%
        group_by(Region=City, Parent=District) %>%
        summarise(Cases=sum(Cases)) %>%
        mutate(Level=3)
    ) %>%
    filter(!is.na(Region)) %>%
    arrange(-Cases) ->
    allSegments
  
  allSegments %>%
    head(numSegment) ->
    topSegments
  
  topSegments %>%
    inner_join(
      topSegments %>%
        transmute(Child=Region, Region=Parent, Level=Level-1, ChildCases=Cases),
      by = c("Region", "Level")
    ) -> 
    kangaroos
  
  topSegmentHistory <- tibble()
  
  iteration <- 1
  
  seenThisBefore <- tibble()
  
  while(nrow(kangaroos) > 0 & iteration < numSegment)  {
    
    kangaroos %>%
      group_by(Region, Cases, Level) %>%
      summarise(ChildCases=sum(ChildCases), Children=paste(Child, collapse=', ')) %>%
      mutate(
        NewRegion=paste0(Region, ' - {', Children, '}'),
        NewCases=Cases-ChildCases
      ) -> 
      newSegments
    
    allSegments %>%
      mutate(Hybrid=F) %>%
      anti_join(
        newSegments %>%
          ungroup() %>%
          select(Region),
        by='Region'
      ) %>%
      bind_rows(
        newSegments %>%
          ungroup() %>%
          select(Region=NewRegion, Cases=NewCases, Level) %>%
          mutate(Hybrid=T)
      ) %>%
      arrange(-Cases) %>%
      head(numSegment) ->
      topSegments
    
    # topSegments %>%
    #   inner_join(topSegmentHistory, by=c("Region")) ->
    #   seenThisBefore
    # 
    # print(seenThisBefore)
    
    if(nrow(seenThisBefore) > 0)
    {
      topSegments
      
      break
    }
    topSegments %>%
      separate(Region, c("Region", "Excluded"), ' - ') %>%
      inner_join(
        topSegments %>%
          transmute(Child=Region, Region=Parent, Level=Level-1, ChildCases=Cases),
        by = c("Region", "Level")
      ) -> 
      kangaroos
    
    topSegments %>%
      select(Region, Cases) %>%
      mutate(Iteration=iteration) %>%
      bind_rows(topSegmentHistory) ->
      topSegmentHistory
    
    iteration <-iteration + 1
    
    print(topSegments)
    print(kangaroos)
    
  }
  
  # Add a few lines of code so that all the exclusions are taken care of properly 
  
  if(nrow(kangaroos) > 0) {
    kangaroos %>%
      group_by(Region, Cases, Level) %>%
      summarise(ChildCases=sum(ChildCases), Children=paste(Child, collapse=', ')) %>%
      mutate(
        NewRegion=paste0(Region, ' - {', Children, '}'),
        NewCases=Cases-ChildCases
      ) -> 
      newSegments
    topSegments %>%
      filter(!Region %in% kangaroos$Region) %>%
      bind_rows(
        newSegments %>%
          select(Region=NewRegion, Cases=NewCases)
      ) ->
      topSegments
  }
  
  
  topSegments %>%
    select(Region, Cases) %>%
    bind_rows(
      tibble(
        Region='Rest',
        Cases=sum(totalCases$Cases) - sum(topSegments$Cases)
      )
    ) %>%
    mutate(Percentage=Cases/sum(Cases)) %>%
    return()
}


```


So we can have a cycling condition. Need to deal with this.

Basically we need ot maintain a history of all the things produced and then cut when something repeats (thefirst thing that repeats gets returned) 




Need to figure out something far simpler. This is leading to mad race conditions

```{r}
segmentation <- function(totalCases, numSegment) {
  
  totalCases %>%
    group_by(Region=State) %>%
    summarise(Cases=sum(Cases)) %>%
    mutate(Level=1, Type='State') %>%
    bind_rows(
      totalCases %>%
        group_by(Region=District, Parent=State) %>%
        summarise(Cases=sum(Cases)) %>%
        mutate(Level=2, Type='District')
    ) %>%
    bind_rows(
      totalCases %>%
        group_by(Region=City, Parent=District) %>%
        summarise(Cases=sum(Cases)) %>%
        mutate(Level=3, Type='City')
    ) %>%
    filter(!is.na(Region)) %>%
    arrange(-Cases) ->
    allSegments
  
  allSegments %>%
    head(numSegment) ->
    topSegments
  
  topSegments %>%
    inner_join(
      topSegments %>%
        transmute(Child=Region, Region=Parent, Level=Level-1, ChildCases=Cases),
      by = c("Region", "Level")
    ) -> 
    kangaroos
  if(nrow(kangaroos) > 0) {
    kangaroos %>%
      group_by(Region, Cases, Level) %>%
      summarise(ChildCases=sum(ChildCases), Children=paste(Child, collapse=', ')) %>%
      mutate(
        NewRegion=paste0(Region, ' - {', Children, '}'),
        NewCases=Cases-ChildCases
      ) -> 
      newSegments
    topSegments %>%
      filter(!Region %in% kangaroos$Region) %>%
      bind_rows(
        newSegments %>%
          ungroup() %>%
          select(Region=NewRegion, Cases=NewCases, Level)
      ) ->
      topSegments
  }
  
  
  topSegments %>%
    ungroup() %>%
    #select(Region, Cases) %>%
    arrange(-Cases) %>%
    bind_rows(
      tibble(
        Region='Rest of India',
        Cases=sum(totalCases$Cases) - sum(topSegments$Cases)
      )
    ) %>%
    mutate(Percentage=Cases/sum(Cases)) %>%
    return()
}



```

```{r}

sgmt <- segmentation(totalCases, 15)
totalCases %>%
  mutate(Index=1:n()) ->
  totalCases

totalCases %>%
  gather(Type, Region, -Cases, -Index)  %>%
  filter(!is.na(Region)) %>%
  inner_join(sgmt %>% mutate(Label=Region) %>% select(Type, Label, Region), by = c("Type", "Region")) ->
  class1

totalCases %>%
  filter(!Index %in% class1$Index) %>%
  gather(Type, Region, -Cases, -Index)  %>%
  filter(!is.na(Region)) %>%
  inner_join(
    sgmt %>%
      arrange(Level) %>%
      fill(Type) %>%
      mutate(Label=Region) %>%
      separate(Region, c("Region", "Excluded"), ' - ') %>%
      select(Region, Type, Label), 
    by = c("Type", "Region")
  ) -> 
  class2

class1 %>%
  distinct(Index, Label) %>%
  bind_rows(
    class2 %>% 
      distinct(Index, Label)
  ) %>%
  right_join(totalCases, by='Index') %>%
  mutate(Label=coalesce(Label, 'Rest of India')) ->
  totalCasesWithLabels

dailyCases %>%
  mutate(State=str_to_title(State), District=str_to_title(District), City=str_to_title(City)) %>%
  inner_join(
    totalCasesWithLabels %>%
      select(State, District, City, Label), 
    by = c("State", "District", "City")
  ) ->
  dailyCasesWithLabels



```


```{r}
labelCaseLines <- function(dailyCases, numSegment) {
  dailyCases %>%
    filter(Cases > 0) %>%
    mutate(State=str_to_title(State), District=str_to_title(District), City=str_to_title(City)) %>%
    group_by(State, District, City) %>%
    summarise(Cases=sum(Cases)) %>%
    ungroup() ->
    totalCases
  
  sgmt <- segmentation(totalCases, numSegment)
  totalCases %>%
    mutate(Index=1:n()) ->
    totalCases
  
  totalCases %>%
    gather(Type, Region, -Cases, -Index)  %>%
    filter(!is.na(Region)) %>%
    inner_join(sgmt %>% mutate(Label=Region) %>% select(Type, Label, Region), by = c("Type", "Region")) ->
    class1
  
  totalCases %>%
    filter(!Index %in% class1$Index) %>%
    gather(Type, Region, -Cases, -Index)  %>%
    filter(!is.na(Region)) %>%
    inner_join(
      sgmt %>%
        arrange(Level) %>%
        fill(Type) %>%
        mutate(Label=Region) %>%
        filter(str_detect(Region, ' - ')) %>%
        separate(Region, c("Region", "Excluded"), ' - ') %>%
        select(Region, Type, Label), 
      by = c("Type", "Region")
    ) -> 
    class2
  
  class1 %>%
    distinct(Index, Label) %>%
    bind_rows(
      class2 %>% 
        distinct(Index, Label)
    ) %>%
    right_join(totalCases, by='Index') %>%
    mutate(Label=coalesce(Label, 'Rest of India')) ->
    totalCasesWithLabels
  
  dailyCases %>%
    mutate(State=str_to_title(State), District=str_to_title(District), City=str_to_title(City)) %>%
    inner_join(
      totalCasesWithLabels %>%
        select(State, District, City, Label), 
      by = c("State", "District", "City")
    ) ->
    dailyCasesWithLabels
  
  return(dailyCasesWithLabels)

}
```

```{r}
labelCaseLines(dailyCases %>% filter(Date >= Sys.Date()-3), 19) %>%
  distinct(State, District, City, Label) %>%
  inner_join(dailyCases, by = c("State", "District", "City")) %>%
  filter(Cases > 0) %>%
  group_by(Date, Label) %>%
  summarise(Cases=sum(Cases)) %>%
  ggplot(aes(x=Date, y=Cases)) + geom_point() + geom_line() + geom_smooth() + facet_wrap(~Label) + theme_bw()
```

```{r}
labelCaseLines(dailyCases %>% filter(Date >= Sys.Date()-3), 7) %>%
  distinct(State, District, City, Label) %>%
  inner_join(dailyCases, by = c("State", "District", "City")) %>%
  filter(Cases > 0) %>%
  group_by(Date, Label) %>%
  summarise(Cases=sum(Cases)) %>%
  ggplot(aes(x=Date, y=Cases, col=Label)) + geom_point(size=0.5)+ geom_smooth(se=F)  + theme_bw() + scale_color_brewer(palette='Set1')
```

```{r}
labelCaseLines(dailyCases %>% filter(Date >= Sys.Date()-3), 11) %>%
  distinct(State, District, City, Label) %>%
  inner_join(dailyCases, by = c("State", "District", "City")) %>%
  group_by(Label, Date) %>%
  summarise(Cases=sum(Cases)) %>%
  arrange(Label, Date) %>%
  group_by(Label) %>%
  mutate(
    Total=cumsum(Cases),
    Growth3=log(Total/lag(Total,3))/3,
    DoublingDays=log(2)/Growth3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingDays[Date==max(Date)]
    ) %>%
  ggplot(aes(x=Date, y=DoublingDays)) + geom_point() + geom_line()  + scale_y_continuous("Doubling Time (number of days)") + theme_minimal()  + labs(title="Covid Cases In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)" ) + scale_x_date('') + facet_wrap(~reorder(paste0(Label, ': ', round(LastDouble), ' days to double'), LastGrowth), scales='free') + theme(strip.text = element_text(face='bold', hjust=0), panel.grid.minor = element_blank(), axis.text = element_text(face='bold'))
  
```

