---
title: "Covid Tracker"
output: html_notebook
---


Buggers at covid19india.org changed their API so need to rework this

```{r}
setwd('~/Documents/work/data work/')
require(tidyverse)
```

```{r}
covidindia <- read_csv('https://api.covid19india.org/csv/latest/raw_data.csv', col_types = cols()) %>%
  mutate(`Age Bracket`=as.numeric(`Age Bracket`))


```

```{r}
rd3 <- read_csv('https://api.covid19india.org/csv/latest/raw_data3.csv', col_types = cols()) %>%
  select(-`Contracted from which Patient (Suspected)`)

```

```{r}
rd4 <- read_csv('https://api.covid19india.org/csv/latest/raw_data4.csv', col_types = cols()) %>%
  select(-`Contracted from which Patient (Suspected)`)
rd5 <- read_csv('https://api.covid19india.org/csv/latest/raw_data5.csv', col_types = cols()) %>%
  select(-`Contracted from which Patient (Suspected)`)
rd6 <- read_csv('https://api.covid19india.org/csv/latest/raw_data6.csv', col_types = cols()) %>%
  select(-`Contracted from which Patient (Suspected)`)
rd7 <- read_csv('https://api.covid19india.org/csv/latest/raw_data7.csv', col_types = cols()) %>%
  select(-`Contracted from which Patient (Suspected)`)


rd3  <- bind_rows(rd3, rd4, rd5, rd6, rd7)
```

```{r}
rd3 %>%
  filter(`Current Status` != "Recovered") %>%
  distinct(`Patient Number`, .keep_all = T) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y')) %>%
  rename(State=`Detected State`, District=`Detected District`, City=`Detected City`) %>%
  group_by(Date, State, District, City) %>%
  summarise(Cases=sum(`Num Cases`))
```

```{r}
covidindia %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y')) %>%
  distinct(`Patient Number`, .keep_all = T) %>%
  rename(State=`Detected State`, District=`Detected District`, City=`Detected City`) %>%
  count(Date, State, District, City, name = 'Cases') %>%
  bind_rows(
    rd3 %>%
      filter(`Current Status` != "Recovered") %>%
      mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y')) %>%
      rename(State=`Detected State`, District=`Detected District`, City=`Detected City`) %>%
      group_by(Date, State, District, City) %>%
      summarise(Cases=sum(`Num Cases`))
  ) %>%
  filter(!is.na(Cases)) ->
  dailyCases
```


```{r}
dailyCases %>%
  pull(Cases) %>%
  sum()
```

```{r}
dailyCases %>%
  group_by(Date) %>%
  summarise(n=sum(Cases)) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n)) %>%
  filter(Cumul > 20) %>%
  ggplot(aes(x=Date, y=Cumul)) + geom_point() + geom_line() + scale_y_log10("Total cases (log scale)") + geom_text(aes(label=Cumul), nudge_y=0.1, size=3, fontface='bold') + xlab("") + ggtitle("Cumulative Growth in number of Covid-19 positive cases in India", subtitle="Logarithmic Y-axis to illustrate exponential growth") + theme_minimal()
```


Show it instead in terms of doublign time

Smooth better using a 3-day CDGR
```{r}
dailyCases %>%
  filter(!is.na(Cases)) %>%
  group_by(Date) %>%
  summarise(n=sum(Cases)) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n)) %>%
  filter(Cumul >= 30) %>%
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3
  ) %>%
  filter(!is.na(DoublingTime3)  & Date < Sys.Date()) %>%
  ggplot(aes(x=Date, y=DoublingTime3)) + geom_point() + geom_line()  + scale_y_continuous("Doubling Time (number of days)", breaks=c()) + theme_minimal() + geom_text(aes(label=round(DoublingTime3,1)), nudge_y=0.2, fontface='bold', size=3) + labs(title="Covid Cases In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)") + scale_x_date('',date_breaks = '2 days', date_labels = '%d-%m') + theme(panel.grid.minor = element_blank(), axis.text = element_text(face='bold'))
```


Let's plot the growth rate as well
```{r}
dailyCases %>%
  group_by(Date) %>%
  summarise(n=sum(Cases)) %>%
  filter(!is.na(Date)) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n)) %>%
  filter(Cumul >= 30) %>%
  mutate(
    Growth1=log(Cumul/lag(Cumul,1)),
    Days=log(2)/Growth1
  ) %>%
  filter(!is.na(Growth1) ) %>%
  ggplot(aes(x=Date, y=Growth1)) + geom_point() + geom_line()  + scale_y_continuous("Daily Growth Rate (%)", breaks=c()) + theme_minimal() + geom_text(aes(label=scales::percent(Growth1, .1)), nudge_y=0.002, fontface='bold', size=3) + ggtitle("Covid Cases In India: Growth Rate") + scale_x_date('',date_breaks = '1 day', date_labels = '%d/%m') + theme(panel.grid.minor = element_blank())
```

```{r}
dailyCases %>%
  group_by(Date, State) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 5 & Total >= 50) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingTime3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) & Date >= as.Date('2020-03-26')) %>%
  mutate(StateCase=paste(State, paste(Total, 'total cases'), paste("Now doubling every", round(LastDouble, 1), "days"),  sep='\n')) %>%
  ggplot(aes(x=Date, y=DoublingTime3)) + geom_point() + geom_line()  + scale_y_continuous("Doubling Time (number of days)", breaks=c()) + theme_minimal() + geom_text(aes(label=round(DoublingTime3,1), y=1.1*DoublingTime3), fontface='bold', size=3) + labs(title="Covid Cases In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)" ) + scale_x_date('') + facet_wrap(~reorder(StateCase, LastGrowth), scales='free') + theme(strip.text = element_text(face='bold', hjust=0), panel.grid.minor = element_blank(), axis.text = element_text(face='bold'))
```

Let's do this by district
```{r}
dailyCases %>%
  group_by(Date, State, District) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  mutate(State=paste0(District, ' (', State, ')')) %>%
  group_by(Date, State) %>%
  summarise(n=sum(n)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 5 & Total >= 150) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingTime3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) & Date==max(Date) & DoublingTime3 < 50 & DoublingTime3 > 0) %>%
  mutate(StateCase=paste0(State, ' (',Total, ')'))%>%
  ggplot(aes(x=reorder(StateCase, LastDouble), y=LastDouble, fill=log(Total), label=round(LastDouble, 1))) + geom_col() + geom_text(size=2.3, aes(y=ifelse(LastDouble < 6, LastDouble + 1, LastDouble - 1), col=ifelse(LastDouble < 6, 'black', 'white')), fontface='bold') + scale_y_continuous("Days to double", breaks=c()) + theme_minimal() + labs(title="Covid Cases By District In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)\n Data source: covid19india.org" ) + xlab('') + coord_flip() + theme(panel.grid.minor = element_blank(), axis.text = element_text(face='bold'), legend.position = 'none') + scale_fill_gradient() + scale_colour_identity() + geom_hline(yintercept = 14, col='red', lwd=0.5)
```

```{r}
dailyCases %>%
  group_by(Date, State, District) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  mutate(State=paste0(District, ' (', State, ')')) %>%
  group_by(Date, State) %>%
  summarise(n=sum(n)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 5 & Total >= 200) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingTime3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) & Date >= Sys.Date() - 10) %>%
  mutate(StateCase=paste(State, paste(Total, 'total cases'), paste("Now doubling every", round(LastDouble, 1), "days"),  sep='\n')) %>%
  ggplot(aes(x=Date, y=DoublingTime3)) + geom_point() + geom_line()  + scale_y_continuous("Doubling Time (number of days)", breaks=c()) + theme_minimal() + geom_text(aes(label=round(DoublingTime3,1), y=1.1*DoublingTime3), fontface='bold', size=3) + labs(title="Covid Cases In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)" ) + scale_x_date('') + facet_wrap(~reorder(StateCase, LastGrowth), scales='free') + theme(strip.text = element_text(face='bold', hjust=0), panel.grid.minor = element_blank(), axis.text = element_text(face='bold'))
```

Do the bar graph only for Karnataka
```{r}
dailyCases %>%
  group_by(Date, State, District) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  filter(State=='Karnataka') %>%
  mutate(State=District) %>%
  group_by(Date, State) %>%
  summarise(n=sum(n)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 3 & Total >= 3) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingTime3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) & Date==max(Date)) %>%
  mutate(StateCase=paste0(State, ' (',Total, ')'))%>%
  ggplot(aes(x=reorder(StateCase, LastDouble), y=LastDouble, fill=log(Total), label=round(LastDouble, 1))) + geom_col() + geom_text(size=2.3, aes(y=ifelse(LastDouble < 6, LastDouble + 2, LastDouble - 2), col=ifelse(LastDouble < 6, 'black', 'white')), fontface='bold') + scale_y_continuous("Days to double", breaks=c()) + theme_minimal() + labs(title="Covid Cases By District In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)\n Data source: covid19india.org" ) + xlab('') + coord_flip() + theme(panel.grid.minor = element_blank(), axis.text = element_text(face='bold'), legend.position = 'none') + scale_fill_gradient() + scale_colour_identity() + geom_hline(yintercept = 14, col='red', lwd=0.5)
```

```{r}
dailyCases %>%
  filter(!is.na(Date) & State %in% c('Uttar Pradesh',"Bihar", "Tamil Nadu", "Gujarat", "Rajasthan", "Madhya Pradesh", "Maharashtra", "Punjab")) %>%
  group_by(Date, State, District) %>%
  summarise(n=sum(Cases)) %>%
  group_by(State) %>%
  complete(Date, District, fill=list(n=0)) %>%
  group_by(State, District) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 3 & Total >= 10) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)],
    LastDouble=DoublingTime3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) & Date==max(Date) & DoublingTime3 <= 50 & DoublingTime3 >0) %>%
  mutate(StateCase=paste0(District, ' (',Total, ')'))%>%
  ggplot(aes(x=reorder(StateCase, LastDouble), y=LastDouble, fill=log(Total), label=round(LastDouble, 1))) + geom_col() + geom_text(size=2.3, aes(y=ifelse(LastDouble < 6, LastDouble + 2, LastDouble - 2), col=ifelse(LastDouble < 6, 'black', 'white')), fontface='bold') + scale_y_continuous("Days to double", breaks=c()) + theme_minimal() + labs(title="Covid Cases By District In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate", caption="© Karthik Shashidhar (@karthiks)\n Data source: covid19india.org" ) + xlab('') + coord_flip() + theme(panel.grid.minor = element_blank(), axis.text = element_text(face='bold'), legend.position = 'none') + scale_fill_gradient() + scale_colour_identity() + facet_wrap(~State, scales='free', nrow=2) + geom_hline(yintercept = 14, col='red', lwd=0.5)
```





Only for Karnataka
```{r}
dailyCases %>%
  filter(State=="Karnataka") %>%
  group_by(Date, State) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 5 & Total >= 50) %>%  # ignore small states or states without much data
  mutate(
    Growth3=log(Cumul/lag(Cumul,3)),
    DoublingTime3=log(2)/Growth3*3,
    LastGrowth=Growth3[Date==max(Date)]
  ) %>%
  filter(!is.na(DoublingTime3) ) %>%
  mutate(StateCase=paste(State, paste(Total, 'total cases'), sep='\n')) %>%
  ggplot(aes(x=Date, y=DoublingTime3)) + geom_point() + geom_line()  + scale_y_continuous("Doubling Time (number of days)", breaks=c()) + theme_minimal() + geom_text(aes(label=round(DoublingTime3,1), y=1.1*DoublingTime3), fontface='bold', size=3) + ggtitle("Covid Cases In India: Days to Double", subtitle = "Calculated using 3-day lagging compounded daily growth rate") + scale_x_date('') + facet_wrap(~reorder(StateCase, LastGrowth), scales='free') + theme(strip.text = element_text(face='bold', hjust=0), panel.grid.minor = element_blank(), axis.text = element_text(face='bold'))
```

State wise absolute
```{r}
dailyCases %>%
  group_by(Date, State) %>%
  summarise(n=sum(Cases)) %>%
  filter(!is.na(Date)) %>%
  group_by(State) %>%
  arrange(Date) %>%
  mutate(Cumul=cumsum(n),Total=sum(n)) %>%
  filter(Cumul >= 5 & Total >= 100) %>% # ignore small states or states without much data
  ggplot(aes(x=Date, y=Cumul)) + geom_point() + geom_line() + scale_y_log10("Number of cases") + theme_minimal() + facet_wrap(~reorder(State, -Total), scales='free') + ggtitle("Covid Cases In India: Growth by state") + scale_x_date('')  + theme(strip.text = element_text(face='bold', hjust=0)) 
```

Let's predict what the number of cases will be 2 weeks from now based on one-week growth rate

```{r}
dailyCases %>%
  group_by(Date, State) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  filter(!is.na(Date)) %>%
  group_by(State) %>%
  arrange(State, Date) %>%
  mutate(
    Cumul=cumsum(n),
    Today=sum(n),
    Lag7=Cumul[Date==(max(Date)-7)],
    Growth7=log(Today/Lag7)/7,
    Lag3=Cumul[Date==(max(Date)-3)],
    Growth3=log(Today/Lag3)/3,
    Lag2=Cumul[Date==(max(Date)-2)],
    Growth2=log(Today/Lag2)/2,
    GrowthRate=pmax(Growth7, Growth3, Growth2),
    Forecast2weeks=Today*exp(14*GrowthRate)  - Today,  # People infected until today would have bee cured in 2 weeks' time,
    Colour=ifelse(GrowthRate <= log(2)/30, 'yellow', ifelse(GrowthRate <= log(2)/7, 'orange', 'red'))
  ) %>%
  ungroup() %>%
  mutate(StateLabel=paste(State, paste(signif(round(Forecast2weeks),3), 'active cases on', format(Sys.Date()+14, '%d-%b')), sep='\n')) %>%
  filter(Cumul >= 5 & Today >= 50) ->  # ignore small states or states without much data
  stateCumul
stateCumul %>%
  ggplot(aes(x=Date, y=Cumul, col=Colour)) + geom_point() + geom_line() + scale_y_log10("Number of cases") + theme_minimal() + geom_text(aes(label=n, y=1.5*Cumul), fontface='bold', size=2) + facet_wrap(~reorder(StateLabel, -Forecast2weeks), scales='free') + labs(title="Covid Cases In India: Growth by state", subtitle="14 day projection based on maximum of 2-day, 3-day and 7-day growth rate in cases", caption="© Karthik Shashidhar (@karthiks)") + scale_x_date('')  + theme(strip.text = element_text(face='bold', hjust=0)) + scale_colour_identity()
```

The above graph was misleading. Let me change it up so that we now have active cases per day, assuming a recovery period of 14 days, and a growth equal to max of 2, 3, 7 day growth
```{r}
dailyCases %>%
  filter(!is.na(Date)) %>%
  group_by(Date, State) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  complete(Date, State, fill=list(n=0)) %>%
  group_by(State) %>%
  arrange(State, Date) %>%
  mutate(
    Cumul=cumsum(n),
    Today=sum(n),
    Lag7=Cumul[Date==(max(Date)-7)],
    Growth7=log(Today/Lag7)/7,
    Lag3=Cumul[Date==(max(Date)-3)],
    Growth3=log(Today/Lag3)/3,
    Lag2=Cumul[Date==(max(Date)-2)],
    Growth2=log(Today/Lag2)/2,
    GrowthRate=pmax(Growth7, Growth3, Growth2),
    Colour=ifelse(GrowthRate <= log(2)/30, 'yellow', ifelse(GrowthRate <= log(2)/7, 'orange', 'red'))
  ) %>%
  select(Date, State, Cumul, GrowthRate, Colour, Today) %>%
  inner_join(
    (.) %>%
      filter(Date==max(Date)) %>%
      select(Date, State, Today, GrowthRate) %>%
      crossing(Projection=0:14) %>%
      mutate(
        ProjectedDate=Date + Projection, 
        ProjectedCases=Today * exp(Projection * GrowthRate),
        Date=ProjectedDate-14
      ) %>%
      select(-GrowthRate, -Today, -Projection),
    by=c("State", "Date")
  ) %>%
  mutate(
    LiveCases=ProjectedCases-Cumul,
    StateLabel=paste(State, paste(signif(round(Today*exp(14*GrowthRate)-Today),3), 'active cases on', format(Sys.Date()+14, '%d-%b')), sep='\n')
  ) %>%
  filter(Cumul >= 5 & Today >= 50) %>%
  ggplot(aes(x=ProjectedDate, y=LiveCases, col=Colour)) + geom_point() + geom_line() + scale_y_continuous("Number of cases", labels=scales::label_number_si()) + theme_minimal() + geom_text(aes(label=round(LiveCases), y=1.1*LiveCases), fontface='bold', size=2) + facet_wrap(~reorder(StateLabel, -Today*exp(GrowthRate*14)), scales='free') + labs(title="Covid Cases In India: Growth by state", subtitle="14 day projection based on maximum of 2-day, 3-day and 7-day growth rate in cases\nAssuming a constant 14-day recovery period", caption="© Karthik Shashidhar (@karthiks)") + scale_x_date('')  + theme(strip.text = element_text(face='bold', hjust=0)) + scale_colour_identity()



```

```{r}
dailyCases %>%
  mutate(Region=coalesce(City, District, State)) %>%
  group_by(Date, Region) %>%
  summarise(Cases=sum(Cases))
``` 


```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y'))  %>%
  select(Date, StateID=`State Patient Number`, City=`Detected City`, District=`Detected District`,Status=`Current Status`, Notes=Notes, Age=`Age Bracket`, Sex=Gender, Transmission=`Type of transmission`, MoreNotes=`Backup Notes`) %>%
  mutate(Notes=str_sub(Notes, 1, 60)) %>%
  filter(!str_detect(str_to_lower(Notes), 'travel') & !str_detect(str_to_lower(Notes), 'return')) %>%
  arrange(desc(Date), District, desc(StateID))
```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y'))  %>%
  count(Date)
```


```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y'))  %>%
  filter(`Detected District`=='Bengaluru Urban') %>%
  select(Date, StateID=`State Patient Number`,Status=`Current Status`, Notes=Notes, Age=`Age Bracket`, Sex=Gender, Transmission=`Type of transmission`, MoreNotes=`Backup Notes`) %>%
  mutate(Notes=str_sub(Notes, 1, 60)) %>%
  arrange(desc(Date), desc(StateID))
```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y')) %>%
  group_by(Notes, `Detected District`) %>%
  summarise(Num=n(), Earliest=min(Date)) %>%
  arrange(-Num) %>%
  filter(Num>=5)
```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y')) %>%
  group_by(Notes) %>%
  summarise(Num=n(), Earliest=min(Date)) %>%
  arrange(-Num) %>%
  filter(Num>=5)
```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(
    Date=as.Date(`Date Announced`, '%d/%m/%Y'),
    MH=str_detect(Notes, 'Mumbai') | str_detect(Notes, 'Maharashtra') | str_detect(Notes, "Maharastra")
  ) %>%
  count(MH, Date) %>%
  group_by(Date) %>%
  summarise(MH=sum(n[MH], na.rm=T)/sum(n, na.rm=T)) %>%
  arrange(desc(Date))
```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(
    Date=as.Date(`Date Announced`, '%d/%m/%Y'),
    Travel=str_detect(str_to_lower(Notes), 'travel') | str_detect(str_to_lower(Notes), 'return')
  ) %>%
  count(Travel, Date) %>%
  group_by(Date) %>%
  summarise(Travel=sum(n[Travel], na.rm=T), Total=sum(n, na.rm=T)) %>%
  mutate(Local=Total-Travel) %>%
  arrange(Date) %>%
  mutate(SmoothLocal=(Local+lag(Local,1)+lag(Local,2)+lag(Local,3)+lag(Local,4))/5) %>%
  ggplot(aes(x=Date, y=Local)) + geom_point() + geom_line() + geom_line(aes(y=SmoothLocal), col='red', lwd=0.5) +  geom_smooth(method='loess', formula='y~x', se=T) + theme_minimal() + xlab('') + ylab("Number of cases discovered") + labs(title='Number of new locally transmitted covid cases in Karnataka by date of discovery', subtitle='Calculated as total cases minus cases with travel history', caption='Prepared by Karthik Shashidhar, based on data from covid19india.org') + theme(title=element_text(size=16, face='bold'), text=element_text(size=12, face='bold'))


```

```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  filter(`Detected District`=='Bengaluru Urban') %>%
  mutate(
    Date=as.Date(`Date Announced`, '%d/%m/%Y'),
    Travel=str_detect(str_to_lower(Notes), 'travel') | str_detect(str_to_lower(Notes), 'return')
  ) %>%
  count(Travel, Date) %>%
  group_by(Date) %>%
  summarise(Travel=sum(n[Travel], na.rm=T), Total=sum(n, na.rm=T)) %>%
  mutate(Local=Total-Travel) %>%
  arrange(Date) %>%
  mutate(SmoothLocal=(Local+lag(Local,1)+lag(Local,2)+lag(Local,3)+lag(Local,4))/5) %>%
  ggplot(aes(x=Date, y=Local)) + geom_point() + geom_line() + geom_line(aes(y=SmoothLocal), col='red', lwd=0.5) + geom_smooth(method='loess', formula='y~x', se=T) + theme_minimal() + xlab('') + ylab("Number of cases discovered") + labs(title='Number of new locally transmitted covid cases in Bangalore by date of discovery', subtitle='Calculated as total cases minus cases with travel history', caption='Prepared by Karthik Shashidhar, based on data from covid19india.org') + theme(title=element_text(size=16, face='bold'), text=element_text(size=12, face='bold'))

```

Let's try and build a graph for Karnataka
```{r}
covidindia %>%
  filter(`Detected State`=='Karnataka') %>%
  bind_rows(
    rd3 %>%
      filter(`Detected State`=='Karnataka' & `Current Status` !='Recovered' & !is.na(`State Patient Number`))
  ) %>%
  mutate(Date=as.Date(`Date Announced`, '%d/%m/%Y'))  %>%
  select(Date, StateID=`State Patient Number`, City=`Detected City`, District=`Detected District`,Status=`Current Status`, Notes=Notes, Age=`Age Bracket`, Sex=Gender, Transmission=`Type of transmission`, MoreNotes=`Backup Notes`) %>%
  mutate(NoteLength=nchar(Notes)) %>%
  arrange(Date) %>%
  distinct(StateID, .keep_all = T)->
  karCases
```

```{r}
karCases %>%
  filter(str_detect(Notes, 'P[0-9]+') | str_detect(Notes, 'P-[0-9]+')) %>%
  mutate(
    Notes=str_replace(Notes, 'P-', 'P'),
    Contact=paste("KA", str_extract(Notes, 'P[0-9]+'), sep='-'),
    MainID=parse_number(str_extract(StateID, '[0-9]+')),
    ContactID=parse_number(str_extract(Contact, '[0-9]+'))
  ) %>%
  filter(MainID > ContactID) %>%
  select(StateID, ContactID=Contact) ->
  contactGraph
```


Let's see how long it has been taking for contacts of cases to be infected

```{r}
contactGraph %>%
  inner_join(
    karCases %>%
      select(StateID, Date1=Date),
    by='StateID'
  ) %>%
  inner_join(
    karCases %>%
      select(ContactID=StateID, Date2=Date),
    by='ContactID'
  ) %>%
  mutate(DateDiff=as.numeric(Date1-Date2, units='days')) %>%
  ggplot(aes(x=DateDiff)) + geom_histogram(binwidth = 1, col='white', lwd=1)  + theme_minimal() + scale_x_continuous('', breaks=seq(0,30,3))
```

```{r}
contactGraph %>%
  inner_join(
    karCases %>%
      select(StateID, Date1=Date),
    by='StateID'
  ) %>%
  inner_join(
    karCases %>%
      select(ContactID=StateID, Date2=Date),
    by='ContactID'
  ) %>%
  mutate(DateDiff=as.numeric(Date1-Date2, units='days')) %>%
  count(ContactID, DateDiff, sort=T) 
```

```{r}
contactGraph %>%
  count(ContactID, sort=T) %>%
  bind_rows(
    karCases %>%
      filter(Date < Sys.Date() -7) %>%
      select(ContactID=StateID) %>%
      mutate(n=0)
    
  ) %>%
  mutate(x=1:n(), y=cumsum(n)) %>%
  ggplot(aes(x=x, y=y)) + geom_point() + geom_line() + theme_minimal() + scale_x_continuous('"Rank" of spreader (by number of people infected)', breaks = seq(0,1800,50)) + scale_y_continuous("Cumulative number of people infected", breaks=seq(0, 1800,50)) + labs(title="Covid-19 Superspreader curve in Karnataka", subtitle='A small number of infected people are responsible for transmitting a large number of infections', caption='Prepared by Karthik Shashidhar, based on data from covid19india.org') + theme(text=element_text(face='bold', size=12), title=element_text(size=16, face='bold'))
```

```{r}
contactGraph %>%
  inner_join(karCases %>% select(StateID, Date), by = "StateID") %>%
  group_by(ContactID) %>%
  summarise(n=n(), Earliest=min(Date), Latest=max(Date)) %>%
  ungroup() %>%
  rename(StateID=ContactID) %>%
  inner_join(karCases %>% mutate(City=coalesce(City, District)) %>%select(StateID, Date,  Notes, City, Age, Sex), by = "StateID") %>%
  mutate(Notes=str_sub(Notes, 1, 60)) %>%
  arrange(-n) %>%
  mutate(CumNum=cumsum(n),Prop=scales::percent(cumsum(n)/sum(n),1), Num=1:n())
```

```{r}
contactGraph %>%
  inner_join(karCases %>% select(StateID, Date), by = "StateID") %>%
  group_by(ContactID) %>%
  summarise(n=n(), Earliest=min(Date), Latest=max(Date)) %>%
  ungroup() %>%
  rename(StateID=ContactID) %>%
  inner_join(karCases %>% mutate(City=coalesce(City, District)) %>%select(StateID, Date,  Notes, City, Age, Sex), by = "StateID") %>%
  mutate(Notes=str_sub(Notes, 1, 60)) %>%
  arrange(-n) %>%
  mutate(CumNum=cumsum(n),Prop=scales::percent(cumsum(n)/sum(n),1), Num=1:n()) %>%
  filter(str_detect(Notes, 'Return') | str_detect(Notes, 'Travel'))
```

```{r}
contactGraph %>%
  filter(ContactID=='KA-P653')

contactGraph %>%
  filter(StateID=='KA-P196')

karCases %>%
  filter(str_detect(Notes, 'P128'))
```

```{r}
latestContacts <- contactGraph %>%
  mutate(Level=1)


i <- 2 

allContacts <- latestContacts

while(nrow(latestContacts) > 0) {
  latestContacts %>%
    select(-Level) %>%
    left_join(
      contactGraph %>%
        select(
          Contact2ID=ContactID,
          ContactID=StateID
        ),
      by = "ContactID"
    ) %>%
    select(-ContactID) %>%
    filter(!is.na(Contact2ID)) %>%
    rename(ContactID=Contact2ID) %>%
    mutate(Level=i) ->
    latestContacts
  
  allContacts %>%
    bind_rows(latestContacts) ->
    allContacts
  i <- i + 1
  print(paste(i, Sys.time()))
  
}


```

```{r}
allContacts %>%
  inner_join(
    karCases %>%
      select(StateID, Date),
    by='StateID'
  ) %>%
  ungroup() %>%
  group_by(StateID=ContactID) %>%
  summarise(n=n()+1, Earliest=min(Date), Latest=max(Date)) %>%
  inner_join(karCases %>% mutate(City=coalesce(City, District)) %>%select(StateID, Notes, City, Age, Sex), by = "StateID") %>%
  mutate(Notes=str_sub(Notes, 1, 80)) %>%
  arrange(-n)
```

```{r}
karCases %>%
  filter(StateID=='KA-P125')
```


Now let's try the Lebanon algorithm 
```{r}

# Let's put 1% as the cutoff
# Any region with more than 1% of cases needs its own category

totalCases <- dailyCases %>%
  summarise(Total=sum(Cases)) %>%
  pull(Total) %>% 
  first()

cutoff <- 0.01

dailyCases %>%
  filter(!is.na(City)) %>%
  group_by(City, District, State) %>%
  summarise(Cases=sum(Cases)) %>%
  mutate(Share=Cases/totalCases) %>%
  filter(Share > cutoff) -> 
  level0

level0 %>%
  arrange(-Cases) %>%
  group_by(State, District) %>%
  summarise(
    Lower=sum(Cases),
    Label=paste(City, collapase=', ')
  ) ->
  level01

dailyCases %>%
  filter(!is.na(District)) %>%
  group_by(District, State) %>%
  summarise(Cases=sum(Cases)) %>%
  left_join(
    level01,
    by=c("State","District")
  ) %>%
  ungroup() %>%
  mutate(
    Lower=ifelse(is.na(Lower), 0, Lower),
    Cases=Cases-Lower,
    District=ifelse(Lower > 0, paste0(District, " - {", Label, "}"), District)
  ) %>%
  mutate(Share=Cases/totalCases) %>%
  filter(Share > cutoff) ->
  level1

level1 %>%
  arrange(-Cases) %>%
  group_by(State) %>%
  summarise(
    Lower=sum(Cases),
    Label=paste(District, collapse=', ')
  ) ->
  level11

dailyCases %>%
  group_by(State) %>%
  summarise(Cases=sum(Cases)) %>%
  left_join(
    level11,
    by=c("State")
  ) %>%
  ungroup() %>%
  mutate(
    Lower=ifelse(is.na(Lower), 0, Lower),
    Cases=Cases-Lower,
    State=ifelse(Lower > 0, paste0(State, " - {", Label, "}"), State)
  ) %>%
  mutate(Share=Cases/totalCases) %>%
  filter(Share > cutoff) ->
  level2

bind_rows(
  level0 %>%
    ungroup() %>%
    mutate(Region=paste0(City, ' (', District, ' ', State, ')')) %>%
    select(Region, Cases, Share),
  level1 %>%
    ungroup() %>%
    mutate(Region=paste0(District, ' (', State, ')')) %>%
    select(Region, Cases, Share), 
  level2 %>%
    ungroup() %>%
    mutate(Region=paste(State, sep=' - ')) %>%
    select(Region, Cases, Share)
) %>%
  arrange(-Share) %>%
  mutate(Cumul=cumsum(Share))


```
Now if we have top 10 or something, then the "rest" should also come into the picture. How do we do this? 

And nothing should come in negatives if it isn't there in positives. So those should be "added back".

Input parameters - how many, and what cumulative proportion should be covered. But that can't work. If i say "10 things accounting for 70%", it depends upon the distribution. Top 10 may not account for 70%. Maybe I should just have it as "top 10" or "top 70%".


Can I do a 2 pass thingy? 


```{r}

```






```{r}
dailyCases %>%
  mutate(State=str_to_title(State), District=str_to_title(District), City=str_to_title(City)) %>%
  group_by(Level1=State, Level2=District, Level3=City) %>%
  summarise(Cases=sum(Cases)) %>%
  filter(Cases > 0) -> 
  caseData

caseData %>%
  arrange(-Cases)

topClusters <- function(caseData, numRows=NA, coverage=NA, levelDiscount=5) {
  
  caseData %>%
    group_by(Region=Level1) %>%
    summarise(Cases=sum(Cases), Successors=length(unique(Level2, na.rm=T))) %>%
    mutate(Level=1) %>%
    bind_rows(
      caseData %>%
        group_by(Region=Level2) %>%
        summarise(Cases=sum(Cases, Successors=length(unique(Level3, na.rm=T)))) %>%
        mutate(Level=2) 
    ) %>%
    bind_rows(
      caseData %>%
        group_by(Region=Level3) %>%
        summarise(Cases=sum(Cases), Successors=1) %>%
        mutate(Level=3) 
    ) %>%
    filter(!is.na(Region)) %>%
    arrange(-Cases) %>%
    mutate(CasesAdjusted=Cases*(levelDiscount^Level)) %>%
    arrange(-CasesAdjusted)
  
  
}

levelDiscount <- 5
```




Let's just look at the Karnataka districts

```{r}
dailyCases %>%
  filter(!is.na(Date)) %>%
  group_by(Date, District, State) %>%
  summarise(n=sum(Cases)) %>%
  ungroup() %>%
  filter(State=="Karnataka") %>%
  ungroup() %>%
  complete(District, Date, fill = list(n=0)) %>%
  arrange(District, Date) %>%
  group_by(District) %>%
  mutate(
    Total=cumsum(n),
    Latest=Total[Date==max(Date)],
    Back1week=Total[Date==(max(Date)-7)],
    Back3days=Total[Date==(max(Date)-3)],
    WeekGrowth=Latest/Back1week-1,
    Growth3=Latest/Back3days-1,
    colour=ifelse(Latest==Back1week, 'No new cases in last 1 week', ifelse((Back1week==0 & Latest > 0) | WeekGrowth > 1 | Growth3 > 0.7, 'Doubled in 1 week OR 70% growth in last 3 days', 'Low growth in last 1 week'))
  ) %>%
  ggplot(aes(x=Date, y=Total, col=colour)) + geom_line(lwd=2) + facet_wrap(~District, scales='free') + theme_minimal() + scale_y_continuous('', breaks= scales::pretty_breaks()) + scale_x_date('') + scale_colour_manual(values = c('red', 'orange', 'yellow')) + theme(strip.text = element_text(face='bold', hjust=0), legend.position = 'top', legend.title = element_blank(), legend.text = element_text(face='bold'), axis.text = element_text(face='bold')) + labs(title="Growth of Covid-19 cases by district in Karnataka", caption='© Karthik Shashidhar\n  @karthiks')
```

Do it on a choropleth

```{r}
require(sf)
kar <- st_read('~/Documents/work/data work/maps/maps/Districts/Census_2011/2011_Dist.shp') %>%
  filter(ST_NM=='Karnataka')
```

```{r}


kar %>%
  mutate(DISTRICT=as.character(DISTRICT)) %>%
  left_join(
    dailyCases %>%
      filter(!is.na(Date)) %>%
      group_by(Date, District, State) %>%
      summarise(n=sum(Cases)) %>%
      ungroup() %>%
      filter(State=="Karnataka") %>%
      ungroup() %>%
      complete(District, Date, fill = list(n=0)) %>%
      arrange(District, Date) %>%
      group_by(District) %>%
      mutate(
        Total=cumsum(n),
        Latest=Total[Date==max(Date)],
        Back1week=Total[Date==(max(Date)-7)],
        Back3days=Total[Date==(max(Date)-3)],
        WeekGrowth=Latest/Back1week-1,
        Growth3=Latest/Back3days-1,
        colour=ifelse(Latest==Back1week, 'No new cases in last 1 week', ifelse((Back1week==0 & Latest > 0) | WeekGrowth > 1 | Growth3 > 0.7, 'Doubled in 1 week OR 70% growth in last 3 days', 'Low growth in last 1 week'))
      ) %>%
      distinct(District, colour) %>%
      left_join(
        tibble(
          District=c('Bagalkote', 'Ballari', 'Belagavi', 'Bengaluru Urban', "Bengaluru Rural", "Kalaburagi", 'Mysuru', 'Tumakuru', 'Vijayapura', 'Shivamogga', 'Chikkamagaluru'),
          DISTRICT=c('Bagalkot', 'Bellary', 'Belgaum', 'Bangalore', 'Bangalore Rural', 'Gulbarga', 'Mysore', 'Tumkur' , 'Bijapur', 'Shimoga', 'Chikmagalur')
        ),
        by='District'
      ) %>%
      mutate(DISTRICT=coalesce(DISTRICT, District)) %>%
      ungroup() %>%
      distinct(DISTRICT, colour),
    by='DISTRICT'
  ) %>%
  mutate(colour=coalesce(colour, 'Zero cases')) %>%
  ggplot() + geom_sf(aes(fill=colour), col='black') + geom_text(data=st_coordinates(st_centroid(kar)) %>% as_tibble() %>% mutate(DISTRICT=kar$DISTRICT), aes(x=X, y=Y, label=DISTRICT), fontface='bold', size=3)+  scale_fill_manual(values = c('red', 'orange', 'yellow', 'darkgreen'))+ theme_minimal() + theme(legend.position = 'right', legend.title = element_blank()) + labs(title="Prevalence and growth of covid-19 cases in Karnataka, by District", subtitle = format(Sys.Date(), '%d %B, %Y'), caption="© Karthik Shashidhar (@karthiks)") + xlab('') + ylab('') 

```

```{r}
read_csv('https://api.covid19india.org/csv/latest/statewise_tested_numbers_data.csv', col_types = cols()) %>%
  mutate(
    Date=as.Date(`Updated On`, '%d/%m/%Y'),
    Tested=as.numeric(`Total Tested`)
  ) %>%
  rename(state=State) ->
  covidtests
```

```{r}
covidtests %>%
  distinct(state, Date, Tested) %>%
  filter(!is.na(Tested)) %>%
  group_by(state) %>%
  arrange(state, Date) %>%
  mutate(
    TestedToday=Tested-lag(Tested, 1),
    DateDiff=as.numeric(Date-lag(Date, 1), units='days')
  ) %>%
  filter(n() > 5 & !is.na(DateDiff) & DateDiff==1) %>%
  ggplot(aes(x=Date, y=Tested, label=Tested)) + geom_point() + geom_line() + geom_text(aes(y=1.1*Tested), fontface='bold', size=2.5) + theme_minimal() + facet_wrap(~state, scales='free') + xlab('') + scale_y_continuous("Cumulative tested", breaks=c())

```

```{r}
covidtests %>%
  distinct(state, Date, Tested) %>%
  filter(!is.na(Tested)) %>%
  group_by(state) %>%
  arrange(state, Date) %>%
  mutate(
    TestedToday=Tested-lag(Tested, 1),
    DateDiff=as.numeric(Date-lag(Date, 1), units='days')
  ) %>%
  filter(n() > 5 & !is.na(DateDiff) & DateDiff==1 & state=='Karnataka')
```

Let's plot testing as well as positive cases
```{r}
covidtests %>%
  distinct(state, Date, Tested) %>%
  inner_join(
    dailyCases %>%
      filter(!is.na(Date)) %>%
      group_by(Date, State) %>%
      summarise(Positive=sum(Cases)) %>%
      ungroup() %>%
      rename(state=State) %>%
      arrange(state, Date) %>%
      group_by(state) %>%
      mutate(Positive=cumsum(Positive)),
    by=c("Date", "state")
  ) %>%
  filter(complete.cases(.)) %>%
  group_by(state) %>%
  filter(n() > 5) %>%
  ggplot(aes(x=Tested, y=Positive, label=format(Date, '%d-%m'))) + geom_point() + geom_smooth(method='loess', formula='y~x') + facet_wrap(~state, scales='free') + theme_minimal()  + xlab("Cumulative people tested") + ylab("Cumulative people found positive")
```


State-wise coloured graph

```{r}
covidtests %>%
  distinct(state, Date, Tested) %>%
  arrange(state, Date) %>%
  filter(!is.na(Tested)) %>%
  group_by(state) %>%
  mutate(
    NewTests=Tested-lag(Tested, 1)
  ) %>%
  filter(!is.na(NewTests) & NewTests > 0) %>%
  mutate(
    DateDiff=as.numeric(Date-lag(Date,1))
  ) %>%
  filter(DateDiff==1) %>%
  filter(n() >= 5) %>%
  inner_join(
    dailyCases %>%
      filter(!is.na(Date)) %>%
      group_by(Date, State) %>%
      summarise(Positive=sum(Cases)) %>%
      ungroup() %>%
      rename(state=State) %>%
      arrange(state, Date) %>%
      group_by(state) %>%
      mutate(Positive=cumsum(Positive)),
    by=c("Date", "state")
  ) %>%
  ungroup() %>%
  filter(complete.cases(.)) %>%
  group_by(state) %>%
  arrange(state, Date) %>%
  mutate(
    TestsPerPositive=Tested/Positive,
    BetterTesting=TestsPerPositive > lag(TestsPerPositive, 1),
    LastBig=as.numeric(Sys.Date()-max(Date[!is.na(BetterTesting) & !BetterTesting]))
  ) %>%
  distinct(state, LastBig) %>%
  ungroup() %>%
  mutate(
    LastBig=ifelse(LastBig==Inf,NA,LastBig),
    LastBig=coalesce(LastBig, max(LastBig, na.rm=T)),
    BlueNess=(LastBig-min(LastBig))/(max(LastBig)-min(LastBig)),
    Colour=rgb(1,0.4+BlueNess*0.6,0.4-BlueNess*0.4)
  )%>%
  distinct(state, Colour) ->
  stateColours


covidtests %>%
  distinct(state, Date, Tested) %>%
  inner_join(
    dailyCases %>%
      filter(!is.na(Date)) %>%
      group_by(Date, State) %>%
      summarise(Positive=sum(Cases)) %>%
      ungroup() %>%
      rename(state=State) %>%
      arrange(state, Date) %>%
      group_by(state) %>%
      mutate(Positive=cumsum(Positive)),
    by=c("Date", "state")
  ) %>%
  filter(complete.cases(.)) %>%
  group_by(state) %>%
  filter(n() > 5) ->
  stateTestCases

stateColours %>%
  filter(state %in% stateTestCases$state) ->
  stateColours

stateGraphs <- list()

for(i in 1:nrow(stateColours)) {
  stateTestCases %>%
    filter(state==stateColours$state[i]) %>%
    ggplot(aes(x=Tested, y=Positive, label=format(Date, '%d-%m'))) + geom_point() + geom_smooth(method='loess', formula='y~x') + theme_minimal()  + scale_x_continuous("Total tests", labels=scales::label_number_si()) + scale_y_continuous("Total Positive Cases") + ggtitle(stateColours$state[i]) + theme(plot.background = element_rect(fill=stateColours$Colour[i]), title=element_text(face='bold', size=16), text=element_text(face='bold', size=10)) ->
    stateGraphs[[i]]
}

require(patchwork)
stateGraphs %>% wrap_plots(nrow=3, guides = 'collect') ->
  p

ggsave('tmp.png', p, width=20, height=10, units='in')





```




```{r}
stateTestCases %>%
  group_by(state) %>%
  mutate(Label=ifelse(Date==max(Date), state, '')) %>%
  inner_join(stateColours, by='state') %>%
  ggplot(aes(x=Tested, y=Positive, group=state, col=Colour, label=Label)) + geom_line(lwd=1) + geom_point() + geom_text(nudge_x=10,fontface='bold', col='black', size=3, hjust=0)+ scale_colour_identity() + theme_minimal() 
```

```{r}
stateTestCases %>%
  group_by(state) %>%
  mutate(
    Ratio=Tested/Positive,
    Label=ifelse(Date==max(Date), round(Ratio), '')
  ) %>%
  ggplot(aes(x=Date, y=Ratio, label=Label)) + geom_point() + geom_line() + geom_text(hjust=1, nudge_y=1, fontface='bold', size=2.5) + facet_wrap(~state, scales='free') + theme_minimal() + xlab('') + ylab("Tests per positive case")+ labs(title="Tests per positive case", caption="© Karthik Shashidhar (@karthiks)\n Data source: covid19india.org" ) + theme(title=element_text(face='bold'), text=element_text(face='bold'))
```


```{r}
stateTestCases %>%
  group_by(state) %>%
  mutate(
    Label=ifelse(Date==max(Date), state, ''),
    Ratio=Tested/Positive
  ) %>%
  inner_join(stateColours, by='state') %>%
  ggplot(aes(x=Date, y=Ratio, group=state, col=Colour, label=Label)) + geom_line(lwd=1) + geom_point() + geom_text(nudge_x=0.1,fontface='bold', col='black', size=3, hjust=0)+ scale_colour_identity() + theme_minimal() 
```

```{r}
p
```


Now let's try multiple Y axes
Data is same
```{r}
covidtests %>%
  distinct(state, Date, Tested) %>%
  inner_join(
    dailyCases %>%
      filter(!is.na(Date)) %>%
      group_by(Date, State) %>%
      summarise(Positive=sum(Cases)) %>%
      ungroup() %>%
      rename(state=State) %>%
      arrange(state, Date) %>%
      group_by(state) %>%
      mutate(Positive=cumsum(Positive)),
    by=c("Date", "state")
  ) %>%
  filter(complete.cases(.) & Date >= as.Date('2020-04-07')) %>%
  group_by(state) %>%
  filter(n() > 5) %>%
  ggplot(aes(x=Date, y=Tested, col='tested')) + geom_line() + geom_point() + geom_point(aes(y=Positive*20, col='Positive')) + geom_line(aes(y=Positive*20, col='Positive')) + scale_y_continuous("Cumulative Tests (Left Axis)", sec.axis = sec_axis(~./20, name='Cumulative Positive (Right Axis)'), labels=scales::label_number_si()) + facet_wrap(~state, scales='free') + scale_colour_manual(values=c('blue', 'red')) + theme_minimal() + theme(axis.line.y.left = element_line(colour='red'), axis.text.y.left = element_text(colour='red', face='bold'), axis.ticks.y.left = element_line(colour='red'), axis.line.y.right = element_line(colour='blue'), axis.text.y.right = element_text(colour='blue', face='bold'), axis.ticks.y.right = element_line(colour='blue'), axis.title.y.right = element_text(colour='blue', face='bold'), axis.title.y.left=element_text(colour='red', face='bold'), legend.position='none', strip.text = element_text(face='bold', size=12, hjust=0), axis.text.x = element_text(face='bold')) + xlab('')
```


