---
title: "Covid Spread Simulator"
output: html_notebook
---

How dangerous is a shopping trip? What happens if you are in an in-person meeting? What happens at a conference? What happens at home? 

Let's simulate all this

```{r}
lambda <- function(timePeriod, p = 0.5) {
  # timePeriod, in hours, is the period of time you need to spend with an infected person in order to have a *p* chance of get infected yourself
  return(- log (1-p)/timePeriod)
}
```

```{r}
transmissionProb <- function(timePeriod, lambda) {
  # When the disease spreads with a parameter lambda, what is the likelihood you'll get infected when you spend *timePeriod* hours with an infected person? 
  return(1 - exp(-lambda * timePeriod))
}
```

```{r}
lambda1 <- lambda(24)
tibble(Time=1:240) %>%
  mutate(Prob=transmissionProb(Time, lambda1)) %>%
  ggplot(aes(x=Time, y=Prob)) + geom_point() + geom_line() + theme_minimal() + scale_y_continuous("Probability of transmission", labels=scales::percent) + ggtitle("Likelihood of getting infected when you spend N hours with an infected person", subtitle = "Assuming that if you spend 24 hours with an infected person, you're 50% likely to fall ill") + scale_x_continuous("Hours", breaks=seq(0,240,6))
```


How safe is it to go to a shop? 

Let's assume there's one shopkeeper. Only one customer can visit the shop at a time, and each customer spends N minutes there. 

This is similar to home delivery! 

```{r}
shopSimulate1 <- function(totalCustomers, initialIncidence, timePerCustomer, lambda, numSimulations = 1000) {
  # *totalCustomers* customers come to the shop each day, only one at a time.
  # initialy, *initialIncidence* (proportion) of them are infected. 
  # They spend a time *timePerCustomer* (in minutes) at the shop, and the disease spreads at rate lambda PER HOUR
  # Initially assume that the shopkeeper is not infected
  # We'll run *numSimulations* simulations of this
  shopkeeper <- F # initially the shopkeeper is not infected
  numInfected <- rep(0, numSimulations)
  shopkeeperInfected <- rep(F, numSimulations)
  initInfected <- runif(totalCustomers) < initialIncidence # For now let's keep this constant across simulations
  ## Update 30 March - we'll change this 
  
  transmissionProb <- 1 - exp(-lambda * timePerCustomer / 60)  # / 60 because lambda is in per hour, timePerCustomer is in per minute
  mcMatrix <- matrix(runif(numSimulations * totalCustomers), nrow=numSimulations) < transmissionProb
  newInfected <- rep(0, numSimulations)
  for(rep in 1:numSimulations) {
    initInfected <- rbinom(totalCustomers, 1, initialIncidence)  # Added so that we can start with a new base each time.
    finalInfected <- initInfected
    shopkeeper <- F
    for(i in 1:totalCustomers) 
      if(mcMatrix[rep, i])   # transmission happens at this step. If not, status quo
        if(finalInfected[i] | shopkeeper) {
          finalInfected[i] <- T
          shopkeeper <- T
        }
    numInfected[rep] <- sum(finalInfected)
    newInfected[rep] <- sum(finalInfected) - sum(initInfected) 
    shopkeeperInfected[rep] <- shopkeeper
  }
  return(
    tibble(
      Sim=1:numSimulations,
      NumInfected=numInfected,
      ExtraInfections=newInfected,
      ShopkeeperInfected=shopkeeperInfected
    )
  )

    
}
```

```{r}
shopSimulate1(1000, 0.5, 2, lambda1, 10000) %>%
  count(ExtraInfections)
```

```{r}
deliveries <- tibble() 
for(prior in c(0.001, 0.002, 0.005, 0.01, 0.02, 0.05, 0.1, 0.2, 0.5)) {
  shopSimulate1(1000, prior, 2, lambda1, 100000) %>%
    count(ExtraInfections) %>%
    mutate(Prior=prior, Proportion=n/sum(n)) %>%
    bind_rows(deliveries) ->
    deliveries
}





```

```{r}
(
deliveries %>%
  ggplot(aes(x=ExtraInfections, y=Proportion)) + geom_col() + facet_wrap(~reorder(scales::percent(Prior, 0.1), Prior)) + scale_x_continuous("Number of additional people infected through delivery", breaks=0:(max(deliveries$ExtraInfections))) + theme_minimal() + scale_y_continuous('', labels=scales::percent) + theme(strip.text = element_text(face='bold')) + ggtitle("Incidence Of Extra Infections Through 2-minute Delivery, \nby base incidence of infection") 
) +
  (
    deliveries %>%
  filter(ExtraInfections > 0) %>%
  ggplot(aes(x=ExtraInfections, y=Proportion)) + geom_col() + facet_wrap(~reorder(scales::percent(Prior, 0.1), Prior)) + scale_x_continuous("Number of additional people infected through delivery", breaks=1:(max(deliveries$ExtraInfections))) + theme_minimal() + scale_y_continuous('', labels=scales::percent) + geom_text(aes( label=scales::percent(Proportion, 0.1)), nudge_y=0.005, fontface='bold', size=3) + theme(strip.text = element_text(face='bold')) + ggtitle("Incidence Of Extra Infections Through 2-minute Delivery, \nby base incidence of infection", subtitle="Excluding case of zero extra infections")
    
  )
```
```{r}
scales::percent(transmissionProb(20/60, lambda1))
```


Next, let's just assume that people arrive and leave by a Poisson process. Anyone inside the store at any given point in time can infect anyone else


```{r}
gc()
```

code for testing

```{r}
N <- 1000
arrivalInterval <- 10
serviceTime <- 5
initialIncidence <- 0.1
numShopkeepers <- 10
```

OK now let's admit up to N people into the shop.

Gap between arrival of two consecutive people is exponential. Service time of each person is also exponential


```{r}
shopSimulate2 <- function(
  N, 
  arrivalInterval,
  serviceTime,
  initialIncidence,
  numShopkeepers=1,  # they stay in the shop all the time
  lambda # the disease transmission rate from an infected to an uninfected
  #shopCapacity = 10000, # the number of people who can be in the shop at one time. If not there will be a queue outside
)
{
  entryTimes <- cumsum(rexp(N, 1/arrivalInterval))
  shoppingTimes <- rexp(N, 1/serviceTime)
  initInfected <- runif(N) < initialIncidence
  shoppers <- tibble(
    Number=1:N, 
    Entry=entryTimes,
    Exit=entryTimes + shoppingTimes,
    PriorInfection=initInfected,
    LaterInfection=initInfected
  ) %>%
    bind_rows(  # adding the shopkeepers
      tibble(
        Number=(N+1):(N + numShopkeepers),
        Entry=0,
        Exit=max(entryTimes + shoppingTimes),
        PriorInfection=F,
        LaterInfection=F
      )
    )
  shoppers %>%
    gather(Activity, Time, -Number, -PriorInfection, -LaterInfection) %>%
    arrange(Time) %>%
    ungroup() %>%
    mutate(
      Change=ifelse(Activity=='Entry', 1, -1),
      PeopleInShop=cumsum(Change) + numShopkeepers,
      Interval=lead(Time,1)-Time,
      numInitInfected=cumsum(ifelse(PriorInfection, Change, 0))
    ) ->
    shopActivity
  shoppers %>%
    mutate(InfectionTime = ifelse(PriorInfection, 0, NA)) ->
    shoppers
  
  infectedInShop <- 0  # number of people currently in the shop who are infected
  
  for(i in 1:nrow(shopActivity)) {
    infectedInShop <- infectedInShop + ifelse(shopActivity$LaterInfection[i], shopActivity$Change[i], 0)
    
    if(infectedInShop > 0)  { # There is at least one infected person in the shop. Else status quo. 
      transProb <- 1 - exp(-lambda * shopActivity$Interval[i] * infectedInShop / 60)  # / 60 since times are in minutes nad lambda is per hour
      
      shoppers %>%
        filter(Entry <= shopActivity$Time[i] & Exit > shopActivity$Time[i] & !LaterInfection) %>% # Shoppers in store at that time, and yet uninfected
        mutate(Infected=runif(n()) < transProb) %>%
        filter(Infected) %>%
        pull(Number) ->
        newlyInfected
      
      shoppers %>%
        mutate(
          LaterInfection=ifelse(Number %in% newlyInfected, T, LaterInfection),
          InfectionTime=ifelse(Number %in% newlyInfected, shopActivity$Time[i], InfectionTime)
          ) -> 
        shoppers
      
      infectedInShop <- infectedInShop + length(newlyInfected)
      
      
      shopActivity %>%
        mutate(LaterInfection=ifelse(Number %in% newlyInfected & Time > shopActivity$Time[i], T, LaterInfection)) ->
        shopActivity
    }
  }
  
  return(list(shoppers, shopActivity))
}
```

```{r}
shopSimulate2(1000, 1, 5, 0.05, 10, lambda1) ->
  sim2

sim2[[1]] %>%
  filter(LaterInfection) %>%
  arrange(InfectionTime) %>%
  mutate(Total=cumsum(LaterInfection)) %>%
  ggplot(aes(x=InfectionTime, y=Total)) + geom_point() + geom_line()



sim21 <- sim2[[1]]
sim22 <- sim2[[2]]

sum(sim21$LaterInfection)

sim21 %>%
  select(Number, Entry, Exit) %>%
  gather(Type, Time, -Number) %>%
  arrange(Time) %>%
  mutate(PeopleInShop=cumsum(ifelse(Type=='Entry', 1, -1))) %>%
  ggplot(aes(x=Time, y=PeopleInShop)) + geom_point() + geom_line()
```

Simulate a 100 times and see
```{r}

sim1 <- tibble()
for(i in 1:100) 
  shopSimulate2(100, 5, 10, 0.05, 10, lambda1) %>%
  .[[1]] %>%
  mutate(Turn=i) %>%
  bind_rows(sim1) ->
  sim1

sim1 %>%
  filter(LaterInfection & InfectionTime > 0) %>%
  arrange(InfectionTime) %>%
  group_by(Turn) %>%
  mutate(Total=cumsum(LaterInfection)) %>%
  ggplot(aes(x=InfectionTime/60, y=Total, group=Turn)) + geom_line(alpha=0.3) + theme_minimal() + xlab('Time (Hours)') + ylab("Number of new infections") + ggtitle("Number of new people infected over time, with a total of 1000 people", "Initially about 50 customers are infected\nCustomers arrive in shop with average gap of 5 minutes and spend 10 minutes in shop\nThere are 10 shopkeepers")

sim1 %>%
  group_by(Turn) %>%
  summarise(NewInfect=sum(LaterInfection-PriorInfection)) %>%
  pull(NewInfect) %>%
  summary()

```

```{r}

sim2 <- tibble()
for(i in 1:100) 
  shopSimulate2(1000, 20, 10, 0.05, 10, lambda1) %>%
  .[[1]] %>%
  mutate(Turn=i) %>%
  bind_rows(sim1) ->
  sim2

sim2 %>%
  filter(LaterInfection & InfectionTime > 0) %>%
  arrange(InfectionTime) %>%
  group_by(Turn) %>%
  mutate(Total=cumsum(LaterInfection)) %>%
  ggplot(aes(x=InfectionTime/60, y=Total, group=Turn)) + geom_line(alpha=0.3) + theme_minimal() + xlab('Time (Hours)') + ylab("Number of new infections") + ggtitle("Number of new people infected over time, with a total of 1000 people", "Initially about 50 customers are infected\nCustomers arrive in shop with average gap of 5 minutes and spend 10 minutes in shop\nThere are 10 shopkeepers") + xlim(0, 80)


sim2 %>%
  group_by(Turn) %>%
  summarise(NewInfect=sum(LaterInfection-PriorInfection)) %>%
  pull(NewInfect) %>%
  summary()
```

```{r}
sim1 <- tibble()
for(i in 1:20) 
  for(prior in seq(0.001, 0.01, 0.05, 0.1, 0.2, 0.5))
  shopSimulate2(100, 5, 10, prior, 1, lambda1) %>%
  .[[1]] %>%
  mutate(Turn=i) %>%
  bind_rows(sim1) ->
  sim1

sim1 %>%
  filter(LaterInfection & InfectionTime > 0) %>%
  arrange(InfectionTime) %>%
  group_by(Turn) %>%
  mutate(Total=cumsum(LaterInfection)) %>%
  ggplot(aes(x=InfectionTime, y=Total, group=Turn)) + geom_point() + geom_line()
```




Next, simulate a company.

Bunch of N people, of whom n are initially infected, spending T minutes with each other

```{r}
meetingSimulate <- function(N, n, T, lambda, numPaths = 1000) {
  # N is total number of people
  # n in initially infected number
  # T is time they spend together (Hours)
  # lambda is rate of  transmission (in hours)
  
  # Let's assume we sample at one minute
  # Here it DOEESN'T MATTER who is infected. Just the totla number matters
  pathLength <- T * 60
  
  numInfected <- rep(n, numPaths)
  infectedTime <- rbind(numInfected, mat.or.vec(pathLength, numPaths))
  
  for(t in 1:pathLength) {
     transProb <- 1 - exp(-lambda * 1 / 60 * numInfected)  # / 60 since times are in minutes nad lambda is per hour
     additionalInfected <- rbinom(numPaths, N-numInfected, transProb)
     numInfected <- numInfected + additionalInfected
     infectedTime[t+1, ] <- numInfected
  }
  row.names(infectedTime) <- (0:pathLength) / 60
  colnames(infectedTime) <- 1:numPaths
  reshape::melt(infectedTime) %>%
    as_tibble() %>%
    rename(Time=X1, Path=X2, Infected=value) %>%
    return()
  #return(infectedTime)
}
```

```{r}
meetingSimulate(10, 1, 8, lambda1, 10000) %>%
  filter(Time==max(Time)) %>%
  mutate(Infected=Infected-1) %>%
  pull(Infected) %>%
  summary()

```

```{r}

meetingSimulate(40, 1, 4, lambda1, 100) %>%
  ggplot(aes(x=Time, y=Infected, group=Path)) + geom_line(alpha=0.2, lwd=0.5) + theme_minimal() + xlab("Time (hours)") + ylab("Infected people") + ggtitle("A 40 person meeting in a closed room over 4 hours", subtitle = "1 infected person at beginning of meeting")
```

Small change in model. Not everyone mgiht come into contact with everyone else each minute. Assume each person will come into contact with m random people each minute
```{r}
conferenceSimulate <- function(N, n, T, lambda, groupSize=N, numPaths = 1000) {
  # N is total number of people
  # n in initially infected number
  # T is time they spend together (Hours)
  # lambda is rate of  transmission (in hours)
  
  # Let's assume we sample at one minute
  # Here it DOEESN'T MATTER who is infected. Just the totla number matters
  pathLength <- T * 60
  
  numInfected <- rep(n, numPaths)
  infectedTime <- rbind(numInfected, mat.or.vec(pathLength, numPaths))
  
  for(t in 1:pathLength) {
     transProb <- 1 - exp(-lambda * 1 / 60 * numInfected * groupSize / N )  # / 60 since times are in minutes nad lambda is per hour; also if you come into contact with groupSize people at a time, number of infected you come in contact with also decreases proportionally 
     additionalInfected <- rbinom(numPaths, N-numInfected, transProb)
     numInfected <- numInfected + additionalInfected
     infectedTime[t+1, ] <- numInfected
  }
  row.names(infectedTime) <- (0:pathLength) / 60
  colnames(infectedTime) <- 1:numPaths
  reshape::melt(infectedTime) %>%
    as_tibble() %>%
    rename(Time=X1, Path=X2, Infected=value) %>%
    return()
  #return(infectedTime)
}
```

```{r}
(conferenceSimulate(100, 1, 8, lambda1, 2, 10000) %>%
  filter(Time==max(Time)) %>%
  mutate(Additional=Infected-1) %>%
  count(Additional) %>%
  mutate(n=n/sum(n))%>%
  ggplot(aes(x=Additional, y=n, label=scales::percent(n,1))) + geom_col() + geom_text(col='white', nudge_y=-0.03, fontface='bold', size=3) + scale_x_continuous("Additional people infected", breaks=0:100, labels=0:100) + theme_minimal() + theme_minimal() + scale_y_continuous('Prbability', breaks=c()) + theme(strip.text = element_text(face='bold')) + ggtitle("Additional infections after a 100-person day-long conference", subtitle="People congregating two at a time")
) +
(conferenceSimulate(100, 1, 8, lambda1, 3, 10000) %>%
  filter(Time==max(Time)) %>%
  mutate(Additional=Infected-1) %>%
  count(Additional) %>%
  mutate(n=n/sum(n))%>%
  ggplot(aes(x=Additional, y=n, label=scales::percent(n,1))) + geom_col() + geom_text(col='white', nudge_y=-0.03, fontface='bold', size=3) + scale_x_continuous("Additional people infected", breaks=0:100, labels=0:100) + theme_minimal() + theme_minimal() + scale_y_continuous('Prbability', breaks=c()) + theme(strip.text = element_text(face='bold')) + ggtitle("Additional infections after a 100-person day-long conference", subtitle="People congregating three at a time")
) +
  (conferenceSimulate(100, 1, 8, lambda1, 5, 10000) %>%
  filter(Time==max(Time)) %>%
  mutate(Additional=Infected-1) %>%
  count(Additional) %>%
  mutate(n=n/sum(n))%>%
  ggplot(aes(x=Additional, y=n, label=scales::percent(n,1))) + geom_col() + geom_text(col='white', nudge_y=-0.03, fontface='bold', size=3) + scale_x_continuous("Additional people infected", breaks=0:100, labels=0:100) + theme_minimal() + theme_minimal() + scale_y_continuous('Prbability', breaks=c()) + theme(strip.text = element_text(face='bold')) + ggtitle("Additional infections after a 100-person day-long conference", subtitle="People congregating five at a time")
) + 
  (conferenceSimulate(100, 1, 8, lambda1, 10, 10000) %>%
  filter(Time==max(Time)) %>%
  mutate(Additional=Infected-1) %>%
  count(Additional) %>%
  mutate(n=n/sum(n))%>%
  ggplot(aes(x=Additional, y=n, label=scales::percent(n,1))) + geom_col() + geom_text(col='white', nudge_y=-0.03, fontface='bold', size=3) + scale_x_continuous("Additional people infected", breaks=0:100, labels=0:100) + theme_minimal() + theme_minimal() + scale_y_continuous('Prbability', breaks=c()) + theme(strip.text = element_text(face='bold')) + ggtitle("Additional infections after a 100-person day-long conference", subtitle="People congregating ten at a time")
)
```
